---
title: "Stan Data Prep"
author: "Kim-Cuong Nguyen"
date: "03/08/2020"
output: md_document
---

Load libraries and data set. The `panel_data` data set was created by merging household panel data with the tissue UPC characteristics.  
```{r}
library(dplyr)
library(tidyverse)

load("panel_data.RData")
```

# Clean data
Filter out unused brand information

```{r}
unused_brands <- c("ALL AMERICAN","ALL BRAND","ALL COKE PRODUCTS","ALL PEPSI PRODUCTS")
panel_data <- panel_data %>% filter(!brand %in% unused_brands)
```

Split dataset for `diet` and `regular` brands
```{r}
# Brands in the table
sort(unique(panel_data$brand))

# Create a column of higher brand level
sevenup <- c("7 UP","7 UP PLUS","7 UP POMEGRANATE","7 UP RETRO","CHERRY 7 UP")
adirondack <- c("ADIRONDACK","ADIRONDACK CLEAR N NATURAL","ADIRONDACK NATURALS")
bigred <- c("BIG RED","BIG PINEAPPLE","BIG PEACH")
canfield <- c("CANFIELD","CANFIELD 50 50")
cape <- c("CAPE COD","CAPE COD DRY")
coke <- c("COKE","BLACK CHERRY VANILLA COKE","COCA COLA","CAFFEINE FREE COKE CLASSIC","CHERRY COKE","COKE BLAK","COKE C2","COKE CLASSIC","COKE WITH LIME","MEXICAN COCA COLA","VANILLA COKE")
diet7up <- c("DIET 7 UP","DIET CHERRY 7 UP")
dietcoke <- c("DIET COKE","CAFFEINE FREE DIET COKE","COKE VANILLA ZERO","COKE CHERRY ZERO","COKE ZERO","DIET BLACK CHERRY VANILLA COK","DIET CHERRY COKE","DIET COKE WITH LIME","DIET COKE PLUS","DIET COKE WITH SPLENDA","DIET COKE WITH LEMON","DIET VANILLA COKE","VANILLA COKE ZERO")
dietdrpepper <- c("DIET DR PEPPER","CAFFEINE FREE DIET DR PEPPER","DIET CHERRY VANILLA DR PEPPER","DIET CHERRY CHOCOLATE DR PEPP","PEPPER FREE")
dietmdew <- c("DIET MOUNTAIN DEW","DIET MOUNTAIN DEW CAFFEINE FR","DIET MOUNTAIN DEW ULTRA VIOLE","DIET MOUNTAIN DEW CODE RED","DIET MTN DEW SUPERNOVA","DIET MTN DEW VOLTAGE","MOUNTAIN DEW MDX SUGAR FREE")
dietpepsi <- c("DIET PEPSI","CAFFEINE FREE DIET PEPSI","DIET PEPSI JAZZ","DIET PEPSI TWIST","DIET PEPSI NFL KICKOFF","DIET PEPSI WITH LIME","DIET PEPSI VANILLA","DIET WILD CHERRY PEPSI","PEPSI MAX","PEPSI MAX CEASE FIRE","PEPSI ONE")
dietrite <- c("DIET RITE","DIET RITE PURE ZERO")
dietsierra <- c("DIET SIERRA MIST","DIET SIERRA MIST CRANBERRY SP","DIET SIERRA MIST FREE","DIET SIERRA MIST RUBY SPLASH","SIERRA MIST FREE","SIERRA MIST FREE UNDERCOVR OR")
dietsprite <- c("DIET SPRITE","SPRITE ZERO")
drpepper <- c("DR PEPPER","CHERRY VANILLA DR PEPPER","DR PEPPER HERITAGE","DR PEPPER TEN","DR PEPPER RED FUSION")
faygo <- c("FAYGO","FAYGO MOON MIST","FAYGO OHANA")
flathead <- c("FLATHEAD LAKE GOURMET SODA","FLATHEAD LAKE MONSTER")
goose <- c("GOOSE ISLAND ORANGE CREAM","GOOSE ISLAND ROOT BEER")
goya <- c("GOYA","GOYA LIGHT","GOYA MALTA")
grafs <- c("GRAFS","GRAFS 50 50","GRANDPA GRAFS")
hansens <- c("HANSENS","HANSENS SIGNATURE")
jolt <- c("JOLT","JOLT BLUE","JOLT RED","JOLT ULTRA")
mello <- c("MELLO YELLO","MELLO YELLO MELON","MELLO YELLO THE ORIGINAL SMOO")
mdew <- c("MOUNTAIN DEW","MOUNTAIN DEW CAFFEINE FREE","MOUNTAIN DEW DISTORTION","MOUNTAIN DEW CODE RED","MOUNTAIN DEW GAME FUEL","MOUNTAIN DEW MDX","MOUNTAIN DEW PITCH BLACK","MOUNTAIN DEW PITCH BLACK 2","MOUNTAIN DEW REVOLUTION","MOUNTAIN DEW THROWBACK","MOUNTAIN DEW SUPERNOVA","MOUNTAIN DEW TYPHOON","MOUNTAIN DEW WHITE OUT","MOUNTAIN DEW VOLTAGE","MT DEW LIVE WIRE")
pepsi <- c("PEPSI","CAFFEINE FREE PEPSI","PEPSI BLUE","PEPSI HOLIDAY SPICE","PEPSI EDGE","PEPSI NFL KICKOFF","PEPSI SUMMER MIX","PEPSI THROWBACK","PEPSI VANILLA","PEPSI TWIST","PEPSI WITH LIME","PEPSI XL","WILD CHERRY PEPSI")
rc <- c("R C","CHERRY R C")
shasta <- c("SHASTA","SHASTA SHORTZ")
sierra <- c("SIERRA MIST","SIERRA MIST CRANBERRY SPLASH","SIERRA MIST NATURAL","SIERRA MIST SQUEEZE","SIERRA MIST RUBY SPLASH","SIERRA MIST UNDERCOVER ORANGE")
sprite <- c("SPRITE","SPRITE REMIX","SPRITE TROPICAL REMIX","SPRITE REMIX ARUBA JAM")
stewarts <- c("STEWARTS","STEWARTS FOUNTAIN CLASSICS")
sunkist <- c("SUNKIST","SUNKIST CITRUS FUSION","SUNKIST SOLAR FUSION")
tava <- c("TAVA BRAZILIAN SAMBA","TAVA MEDITERRANEAN FIESTA","TAVA TAHITIAN TAMURE")
vault <- c("VAULT","VAULT RED BLITZ")

panel_data <- panel_data %>% mutate(brand=ifelse(brand %in% sevenup,"7 UP",ifelse(brand %in% adirondack,"ADIRONDACK",ifelse(brand %in% bigred,"BIG RED",ifelse(brand %in% canfield,"CANFIELD",ifelse(brand %in% cape,"CAPE COD",ifelse(brand %in% coke,"COKE",ifelse(brand %in% diet7up,"DIET 7 UP",ifelse(brand %in% dietcoke,"DIET COKE",ifelse(brand %in% dietdrpepper,"DIET DR PEPPER",ifelse(brand %in% dietmdew,"DIET MOUNTAIN DEW",ifelse(brand %in% dietpepsi,"DIET PEPSI",ifelse(brand %in% dietrite,"DIET RITE",ifelse(brand %in% dietsierra,"DIET SIERRA MIST",ifelse(brand %in% dietsprite,"DIET SPRITE",ifelse(brand %in% drpepper,"DR PEPPER",ifelse(brand %in% faygo,"FAYGO",ifelse(brand %in% flathead,"FLATHEAD LAKE",ifelse(brand %in% goose,"GOOSE ISLAND",ifelse(brand %in% goya,"GOYA",ifelse(brand %in% grafs,"GRAFS",ifelse(brand %in% hansens,"HANSENS",ifelse(brand %in% jolt,"JOLT",ifelse(brand %in% mello,"MELLO YELLO",ifelse(brand %in% mdew,"MOUNTAIN DEW",ifelse(brand %in% pepsi,"PEPSI",ifelse(brand %in% rc,"RC",ifelse(brand %in% shasta,"SHASTA",ifelse(brand %in% sierra,"SIERRA MIST",ifelse(brand %in% sprite,"SPRITE",ifelse(brand %in% stewarts,"STEWARTS",ifelse(brand %in% sunkist,"SUNKIST",ifelse(brand %in% tava,"TAVA",ifelse(brand %in% vault,"VAULT",brand))))))))))))))))))))))))))))))))))

table(is.na(panel_data$brand)) # All values are converted.

sort(unique(panel_data$brand)) # Re-check all brands

# Create separate dataset for diet & regular brands
dietbrands <- c("AIRFORCE NUTRISODA","DIET 7 UP","DIET BARQS","DIET CITRUS BLAST","DIET COKE","DIET DR PEPPER","DIET GRANDPA GRAFS","DIET HANSENS","DIET MELLO YELLO","DIET MINUTE MAID","DIET MOUNTAIN DEW","DIET MUG","DIET PEPSI","DIET R C","DIET RITE","DIET SCHWEPPES","DIET SIERRA MIST","DIET SLICE","DIET SPRITE","DIET SQUIRT","DIET STEAZ","DIET SUN DROP","DIET SUNKIST","DIET TROPICANA TWISTER","DIET VERNORS","FANTA ZERO ORANGE","JONES ZILCH","MELLO YELLO ZERO","MR PIBB ZERO","VAULT ZERO")

panel_data_diet <- panel_data %>% filter(brand %in% dietbrands)
panel_data_regular <- setdiff(panel_data,panel_data_diet)
```

# DIET DATA

Seperate out people's choices, so each purchase is one unit. The `panel` data separates out household's purchases so that each purchase is only 1 unit. The `panel_data` file has the units aggregated. 
```{r}
#testing function
panel_data_diet %>% filter(panel_id==1100016, upc=="41200000686", week==1153)

panel_data_diet %>% filter(panel_id==1100016, upc=="41200000686", week==1153) %>% uncount(units)

panel_diet <- panel_data_diet %>% uncount(units)
```

Create table 1: the y table, with columns panel id, week, upc purchased, and total number of choices for each panelist. 
```{r}
y_table_diet <- panel_diet %>% select(panel_id, week, upc)

y_table_diet <- y_table_diet %>% group_by(panel_id) %>% mutate(n_choices=n())
```

The y_table looks like:
```{r}
head(y_table_diet)
```

Create table 2: the x table, with columns week, upc, features (including average price), brand, and upc rank in week. 

Find average price for each week. The `weekly_prices` data contains the average weekly price for each UPC (averaged price across all UPC purchases in a given week). The `weekly_prices` data is then merged with our `panel_data_m` so that we have the averaged prices for each UPC. 
```{r}
panel_data_diet <- panel_data_diet %>% mutate(avg_price=dollars/units)

weekly_prices_diet <- panel_data_diet %>% group_by(week, upc) %>% summarise(avg_weekly_price=round(mean(avg_price), 2))

panel_data_diet <- panel_data_diet %>% left_join(weekly_prices_diet, by=c("week", "upc"))
```

For each week, find the top 100 upcs (or less in most weeks) in that week. 

Testing for loop. The files in this chunk are simply testing files to make sure the code in the loop works correctly. 
```{r}
top_upc_week1 <- panel_data_diet %>% filter(week==1129) %>% group_by(upc) %>% summarise(total_units=sum(units)) %>% arrange(desc(total_units)) %>% slice(1:100) %>% select(upc) %>% pull()

units <- panel_data_diet %>% filter(week==1129) %>% group_by(upc) %>% summarise(total_units=sum(units)) %>% arrange(desc(total_units)) %>% slice(1:100) %>% mutate(week_rank=seq(1,n()))

test <- panel_data_diet %>% select(upc, week, avg_weekly_price, small_category, parent_co, vendor, brand, upc_desc, vol_eq, product_type, package, flavor, calorie, caffeine, sweetener_type, juice_percentage) %>% filter(week==1129, upc %in% top_upc_week1) %>% distinct()

test <- test %>% left_join(units, by="upc")

rm(units)

```

Create loop. `weeks` is a vector of all the weeks in the panel data. `top_upc_week` finds the top 100 (or less) UPCs in the filtered week. `units` adds a variable for UPC rank in a given week. `temp` filters the panel data for the week and top UPCs, selects the important characteristics, and then is binded into the `x_table`. The `x_table` is our product choices available in each week for the consumer panel. 
```{r}
weeks_diet <- panel_data_diet %>% distinct(week) %>% pull

x_table_diet <- tibble()

for (i in weeks_diet) {
  top_upc_week <-  panel_data_diet %>% filter(week==i) %>% group_by(upc) %>% summarise(total_units=sum(units)) %>% arrange(desc(total_units)) %>% slice(1:100) %>% select(upc) %>% pull()
  
  units <- panel_data_diet %>% filter(week==i) %>% group_by(upc) %>% summarise(total_units=sum(units)) %>% arrange(desc(total_units)) %>% slice(1:100) %>% mutate(week_rank=seq(1,n()))
  
  temp <- panel_data_diet %>% select(upc, week, avg_weekly_price, small_category, parent_co, vendor, brand, upc_desc, vol_eq, product_type, package, flavor, calorie, caffeine, sweetener_type, juice_percentage) %>% filter(week==i, upc %in% top_upc_week) %>% distinct()
  
  temp <- temp %>% left_join(units, by="upc")
  
  x_table_diet <- x_table_diet %>% bind_rows(temp)
  
  rm(top_upc_week, temp, units)
}

#check that all weeks are included 
length(unique(x_table_diet$week))
length(weeks_diet)
```

The x_table looks like:
```{r}
head(x_table_diet)
```

Summary statistics: 

The `y_table_diet` has 1,196,945 rows, 492 unique UPCs, 573 unique weeks, and 10,043 unique panelists. 
```{r}
length(unique(y_table_diet$upc))

length(unique(y_table_diet$week))

length(unique(y_table_diet$panel_id))
```

The `x_table_diet` has 55,133 rows, 573 unique weeks, and 472 unique UPCs. On average, each week has about 96 UPCs (less than 100). 
```{r}
length(unique(x_table_diet$upc))

length(unique(x_table_diet$week))
```

Save the two tables. 
```{r}
save(y_table_diet, file="y_table_diet.RData")
save(x_table_diet, file="x_table_diet.RData")
```


# REGULAR DATA

Seperate out people's choices, so each purchase is one unit. The `panel` data separates out household's purchases so that each purchase is only 1 unit. The `panel_data` file has the units aggregated. 
```{r}
panel_regular <- panel_data_regular %>% uncount(units)
```

Create table 1: the y table, with columns panel id, week, upc purchased, and total number of choices for each panelist. 
```{r}
y_table_regular <- panel_regular %>% select(panel_id, week, upc)

y_table_regular <- y_table_regular %>% group_by(panel_id) %>% mutate(n_choices=n())
```

The y_table looks like:
```{r}
head(y_table_regular)
```

Create table 2: the x table, with columns week, upc, features (including average price), brand, and upc rank in week. 

Find average price for each week. The `weekly_prices` data contains the average weekly price for each UPC (averaged price across all UPC purchases in a given week). The `weekly_prices` data is then merged with our `panel_data_m` so that we have the averaged prices for each UPC. 
```{r}
panel_data_regular <- panel_data_regular %>% mutate(avg_price=dollars/units)

weekly_prices_regular <- panel_data_regular %>% group_by(week, upc) %>% summarise(avg_weekly_price=round(mean(avg_price), 2))

panel_data_regular <- panel_data_regular %>% left_join(weekly_prices_regular, by=c("week", "upc"))
```

For each week, find the top 100 upcs (or less in most weeks) in that week. 

Create loop. `weeks` is a vector of all the weeks in the panel data. `top_upc_week` finds the top 100 (or less) UPCs in the filtered week. `units` adds a variable for UPC rank in a given week. `temp` filters the panel data for the week and top UPCs, selects the important characteristics, and then is binded into the `x_table`. The `x_table` is our product choices available in each week for the consumer panel. 
```{r}
weeks_regular <- panel_data_regular %>% distinct(week) %>% pull

x_table_regular <- tibble()

for (i in weeks_regular) {
  top_upc_week <-  panel_data_regular %>% filter(week==i) %>% group_by(upc) %>% summarise(total_units=sum(units)) %>% arrange(desc(total_units)) %>% slice(1:100) %>% select(upc) %>% pull()
  
  units <- panel_data_regular %>% filter(week==i) %>% group_by(upc) %>% summarise(total_units=sum(units)) %>% arrange(desc(total_units)) %>% slice(1:100) %>% mutate(week_rank=seq(1,n()))
  
  temp <- panel_data_regular %>% select(upc, week, avg_weekly_price, small_category, parent_co, vendor, brand, upc_desc, vol_eq, product_type, package, flavor, calorie, caffeine, sweetener_type, juice_percentage) %>% filter(week==i, upc %in% top_upc_week) %>% distinct()
  
  temp <- temp %>% left_join(units, by="upc")
  
  x_table_regular <- x_table_regular %>% bind_rows(temp)
  
  rm(top_upc_week, temp, units)
}

#check that all weeks are included 
length(unique(x_table_regular$week))
length(weeks_regular)
```

The x_table looks like:
```{r}
head(x_table_regular)
```

Summary statistics: 

The `y_table_regular` has 2,112,095 rows, 3,014 unique UPCs, 573 unique weeks, and 11,545 unique panelists. 
```{r}
length(unique(y_table_regular$upc))

length(unique(y_table_regular$week))

length(unique(y_table_regular$panel_id))
```

The `x_table_regular` has 57,300 rows, 573 unique weeks, and 1,433 unique UPCs. On average, each week has 100 UPCs. 
```{r}
length(unique(x_table_regular$upc))

length(unique(x_table_regular$week))
```

Save the two tables. 
```{r}
save(y_table_regular, file="y_table_regular.RData")
save(x_table_regular, file="x_table_regular.RData")
```
